Automatic Visual Inference for Linear Regression

### 2022-03-16 Milestone_presentation_1/Milestone_presentation_1/Milestone_presentation_1.Rmd
add more inc, fix first part of slides
finish human subject

---
title: "Automatic Visual Inference for Linear Regression"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      beforeInit: "scripts/my_macro.js"
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

@@

#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot. Under the null hypothesis $X ∼ Binom(K, 1/m)$. The p-value of a lineup of size $m$ evaluated by $K$ observer is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i},$$ where $x$ is the realization of number of observers correctly picking the actual data plot.

@@

@@

@@

#### The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

@@

.caption[
Fig. 8: _A lineup of residual plots generated with the cubic model._ 
]
]

@@

.caption[
Fig. 9: _Residual plots of cubic model under different effect sizes._ 
]
]

@@

#### $\boldsymbol{X}$ could be one of the five distributions mentioned above - normal distribution, uniform distribution, adjusted log-normal distribution, adjusted negative log-normal distribution and discrete uniform distribution.

@@

.caption[
Fig. 10: _A lineup of residual plots generated with the heteroskedasticity model._ 
]
]

@@

.caption[
Fig. 11: _Residual plots of heteroskedasticity model under different effect sizes._ 
]

@@

.caption[
Fig. 12: _Layout of the study webiste._ 
]
]

@@

#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes. **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.

#### The same lineup will not be presented to a participant twice. Most of the lineups have been evaluated by **5** different participants.

@@

# $p$-value and Power Estimation for Multiple Selections

#### The assumptions about the **independence** still needed. Let $K$ be the number of independent evaluations of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections. Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$

#### The distribution of $X$ given $s_1, ...s_K$ can not be **derived trivially**, as it is a **sampling without replacement problem**. In practice, this distribution can be **approximated** by computer simulation.

@@

# $p$-value and Power Estimation for Multiple Selections

#### Assume there is a **visual test** $V_K$, where $K$ denoting the number of evaluations. The corresponding  $p$-value of $V_K$ can be computed.

#### Meanwhile, if one evaluation is **randomly deleted** from $V_K$, the remaining evaluations can still be used to form another valid visual test $V_{K-1}$.

#### In fact, considering **all the possibilities**, $K$ different outcomes for $V_{K-1}$ can be obtained.

@@

#### Similarly, if we would like to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$, we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$. Then, the estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.

@@

#### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II. In total, **442** unique lineups are evaluated by 64 subjects.

@@

#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with natural logarithm of effect size as the only regressor.

#### Another method we use to model the power is to collect all decisions made by $K\choose{K-j}$ different outcomes for $V_{K-j}$, and directly fit **binomial generalized linear models** with natural logarithm of effect size as the regressor.

@@

.caption[
Fig. 13: _Quasibinomial GLM fitted for estimated power for different number of evaluations. The left plot is for the cubic model, the right plot is for the heteroskedasticity model. Power curves for different number of evaluations are drawn in different colors. The dashed line is the power of the conventional test._ 
]]

@@

.caption[
Fig. 13: _Binomial GLM fitted for rejection for different number of evaluations. The left plot is for the cubic model, the right plot is for the heteroskedasticity model. Power curves for different number of evaluations are drawn in different colors. The dashed line is the power of the conventional test._ 
]]

@@

#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots, with a focus on the specification error of classical normal linear regression.

@@

#### The system is built upon **computer vision models**. 
]

>>>>

---
title: "Automatic Visual Inference for Linear Regression"
subtitle: ""  
author: 
  - "Weihao (Patrick) Li"
date: '`r format(Sys.Date(), format="%B %d, %Y")`'
output:
  xaringan::moon_reader:
    seal: false
    css: [xaringan-themer.css, mine.css]
    nature:
      beforeInit: "scripts/my_macro.js"
      countIncrementalSlides: false
      slideNumberFormat: "%current%/%total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
---

@@

#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot. Under the null hypothesis $X ∼ Binom(K, 1/m)$.

#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot. Under the null hypothesis $X ∼ Binom(K, 1/m)$.

#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$

#### If we involve $K$ independent observers in a visual test, and let $X$ be a random variable denoting the number of observers correctly picking the actual data plot. Under the null hypothesis $X ∼ Binom(K, 1/m)$.

#### The p-value is given as $$P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}.$$

@@

exclude: true

@@

--

@@

--

#### Under the null hypothesis, $\boldsymbol{e} \sim N(\boldsymbol{0}, \sigma^2\boldsymbol{R}_a)$. The **effect size** is derived as: $$ES=n^{-1}||\sigma^{-1}\boldsymbol{R}_a^{-\frac{1}{2}}\boldsymbol{R}_a\boldsymbol{X}_b\boldsymbol{\beta}_b||^2=n^{-1}\sigma^{-2}||\boldsymbol{R}_a^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2\approx n^{-1}\sigma^{-2}||diag(\boldsymbol{R}_a)^{\frac{1}{2}}\boldsymbol{X}_b\boldsymbol{\beta}_b||^2,$$ where $diag(\boldsymbol{R}_a)$ is the diagonal matrix constructed from the diagonal elements of $\boldsymbol{R}_a$.

@@

.caption[
Fig. 5: _A lineup of residual plots generated with the cubic model._ 
]
]

@@

.caption[
Fig. 6: _Residual plots of cubic model under different effect sizes._ 
]
]

@@

--

#### $X_i$, $i=1,...,n$ follows one of the five distributions:
- $N(0, 0.09)$
- $U(-1, 1)$
- $Lognormal(0,0.36)/3 - 1$
- $-Lognormal(0,0.36)/3 + 1$
- Discrete uniform distribution

@@

.caption[
Fig. 7: _A lineup of residual plots generated with the heteroskedasticity model._ 
]
]

@@

.caption[
Fig. 8: _Residual plots of heteroskedasticity model under different effect sizes._ 
]

@@

.caption[
Fig. 9: _Layout of the study webiste._ 
]
]

@@

.caption[
Fig. 9: _Layout of the study webiste._ 
]
]

#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.

.caption[
Fig. 9: _Layout of the study webiste._ 
]
]

#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.

#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.

.caption[
Fig. 9: _Layout of the study webiste._ 
]
]

#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.

#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.

#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.

.caption[
Fig. 9: _Layout of the study webiste._ 
]
]

#### **20** lineups are given to each participant, with **10** cubic lineups and **10** heteroskedasticity lineups across different effect sizes.

#### **2** out of **20** lineups are extremely easy lineup which are used as attention checks. Participants need to pass at least one.

#### The same lineup will not be presented to a participant twice. Most of the lineups have **5** replicates.

@@

# $p$-value Estimation for Multiple Selections

#### Let $K$ be the number of **independent evaluations** of a visual test, $s_i , i = 1, ..., K$ be the **number of selections of the evaluations**, and $X$ be the random variable denoting the number of detections. Then, the $p$-value of the visual test is given as: $$P(X\geq x) = \sum_{j=x}^{K}Pr(j|s_1,...,s_K).$$

--

#### The distribution of $X$ given $s_1, ...s_K$ can not be **derived trivially**, as it is a **sampling without replacement problem**.

--

#### In practice, this distribution can be **approximated** by computer simulation.

@@

# Power Estimation for Multiple Selections

#### Suppose there is a **visual test** $V_K$, where $K$ denoting the number of evaluations.

--

#### If one evaluation is **randomly deleted** from $V_K$, the **remaining evaluations** can still be used to form **another valid visual test** $V_{K-1}$.

--

#### Considering **all the possibilities**, $K$ different outcomes of $V_{K-1}$ can be obtained.

--

@@

# Power Estimation for Multiple Selections

#### Similarly, to estimate the power of the visual test $V_{K-j}$ given the evaluations of $V_{K}$, for $j < K$,

--

#### we could find all the possible combinations of $K$ elements, taken $k-j$ at a time to obtain $K\choose{K-j}$ different outcomes for $V_{K-j}$.

--

#### The estimated power is given as $R/{K\choose{K-j}}$, where $R$ is the number of outcomes which reject the null hypothesis.

@@

#### We collect **400** lineup evaluations made by **20** participants in experiment I and **880** lineup evaluations made by **44** participants in experiment II.

--

#### In total, **442** unique lineups are evaluated by **64** subjects.

--

@@

--

#### The estimated power is further used in fitting **quasi-binomial generalized linear models** with **natural logarithm of effect size** as the only regressor.

--

#### All **decisions** made by $K\choose{K-j}$ different outcomes of $V_{K-j}$ are fitted by **binomial generalized linear models** with **natural logarithm of effect size** as the regressor.

@@

.caption[
Fig. 10: _Quasibinomial GLM fitted for estimated power for different number of evaluations. The left plot is for the cubic model, the right plot is for the heteroskedasticity model. Power curves for different number of evaluations are drawn in different colors. The dashed line is the power of the conventional test._ 
]]

@@

.caption[
Fig. 11: _Binomial GLM fitted for rejection for different number of evaluations. The left plot is for the cubic model, the right plot is for the heteroskedasticity model. Power curves for different number of evaluations are drawn in different colors. The dashed line is the power of the conventional test._ 
]]

#### There is a huge gap between the conventional curves and the visual test curves at the middle of the plot.

#### Human is hard to identify the pattern at this level of difficulty.

.pull-left[
![](images/mid1.png)

.pull-right[
![](images/mid2.png)

.center[.caption[
Fig. 12: _A cubic lineup with effect size $= 0.18$ and a heteroskedasticity lineup with effect size $= 17.5$._ 
]]

@@

#### We proposed an **automatic visual inference system** for evaluating lineups of residual plots of classical normal linear regression.

@@
